---
title: "Diamonds-Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(ggplot2)
library(tidyverse)
library(data.table)
library(plotly)
library(fastDummies)
library(reshape2)
library(metan)
require(caTools)
library(rpart)
library(gensvm)
library(caret)
library(outliers)
```

## EDA

The aim of this study is to examine the relationships and distributions of variables using the Diamonds dataset available in R and finally develop a meaningful price estimation model. You can access and review the libraries used in this study from the Libraries section.

```{r}
data <- diamonds
```

## Including Plots

There are 53940 rows and 10 columns in diamonds dataset

```{r}
dim(data)
```

```{r}
head(data)
```

```{r}
data %>%glimpse()
```
We can see all data summary like median, mean and quartiles

```{r}
summary(data)
```
There is no missing values

```{r}
sum(is.na(data))
```
we can see table variable's distribution, we see that the concentration is mostly around 50-60.

```{r}
ggplot(data) +
  aes(x = table) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()
```
We can examine the normal and logarithmic normal distribution of our target variable.

```{r}
par(mfrow=c(1,2))
qqnorm((diamonds$price),main="Normal Q-Q Plot of Price");qqline((diamonds$price))
qqnorm(log(diamonds$price),main="Normal Q-Q Plot of log Price");qqline(log(diamonds$price))
```
When we examine the x, y and z variables, we see that the average is at Fair level in the cut variable

```{r}
data %>%
  select(cut,price,x,y,z)%>%
  group_by(cut)%>%
  summarise(mean_price = mean(price),mean_x = mean(x),mean_y = mean(y),mean_z = mean(z))%>%
  arrange(mean_price,desc())
```
When we want to examine the variables with the help of boxplot, we observe that there are outlier values.

```{r}
data%>%
  select(x,cut)%>%
  ggplot(aes(x,color = cut))+
  geom_boxplot()
  
data%>%
  select(y,cut)%>%
  ggplot(aes(y,color = cut))+
  geom_boxplot()

data %>%
  ggplot(aes(x=cut,y=price, color=cut)) +
  geom_boxplot()
```

When the relationship between carat and price is examined, we can say that the price variable also increases as Premium increases.

```{r}
d <- diamonds[sample(nrow(diamonds), 1000), ]

fig <- plot_ly(
  d, x = ~carat, y = ~price,
  # Hover text:
  text = ~paste("Price: ", price, '$<br>Cut:', cut),
  color = ~carat, size = ~carat
)

fig
```

```{r}
data %>%
  group_by(cut) %>%
  summarise(n=n(), 
            mean= mean(price), 
            median=median(price), 
            Q1= quantile(price,0.25),
            Q3= quantile(price,0.75))
```

```{r}
df <- data[,-c(2,3,4)]
df <- rm.outlier(df, fill = TRUE, median = TRUE, opposite = FALSE)
data[,colnames(df)] <- df
head(data)
```
We assign dummies to categorical variables in order to include categorical variables in the model.

```{r}
data <- dummy_cols(data, select_columns = c("cut","color","clarity"),remove_selected_columns = TRUE)
```

```{r}
set.seed(123)
split = sample.split(data$price, SplitRatio = 0.8)
training_set = subset(data, split == TRUE)
test_set = subset(data, split == FALSE)
```

```{r}
cormat <- round(cor(training_set),2)

melted_cormat <- melt(cormat)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

upper_tri <- get_upper_tri(cormat)

cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

print(ggheatmap)

ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 2) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))

```

We are scaling on the train and test data that we finally reached. After the scale result is regression run, we leave only the significant variables and run the model again.

```{r}
training_set = scale(training_set)
test_set = scale(test_set)

d1 <- data.frame(training_set)
d2 <- data.frame(test_set)

regressor = lm(formula = price ~ .,
               data = d1)


summary(regressor)
```

```{r}
d1 <- d1[,-c(7,12,19,27)]
d2 <- d2[,-c(7,12,19,27)]

regressor = lm(formula = price ~ .,
               data = d1)

summary(regressor)

pred <- predict(regressor, newdata = d2)

RMSE(pred = pred, obs = d2$price)
```

# References

http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization
